Equalizer Programming Guide

### 2007-11-27 Developer/ProgrammingGuide/paper.tex
Precursory work to support RenderContext association for window events. Remove glX event thread and implement message pumping for glx. Update doc accordingly.
Needs make clean.

[x] May break build
[ ] Breaks existing applications
[x] Bugfix
[ ] New Feature
[x] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.1.3 for Equalizer 0.4+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.0.1   & Nov 06, 2007 & Fixed Typo\\
    1.1     & Nov 12, 2007 & Adapted distributed objects\\
    1.1.1   & Nov 16, 2007 & Minor improvements based on lulu.com print\\
    1.1.2   & Nov 19, 2007 & Minor improvements based on feedback\\
    1.1.3   & Nov 23, 2007 & Add OpenGL extension handling\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

Where possible events are received and processed by a separate per-node
event thread to allow asynchronous\footnote{with respect to the
  rendering} event handling. Currently an event thread is only used by
the X11/glX window system. WGL receives and processes the events from
the pipe threads that created the windows. AGL receives the events from
application or node main thread. Whenever the term \textbf{event thread}
is used, it refers to the thread receiving the event, such as a per-node
thread for glX, the pipe thread for WGL and the main thread for AGL.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/eventInit.pdf}
  {\caption{\small\label{fEventInit}Event Handling Initialization}}
\end{wrapfigure}
An event handler consists of two parts: the generic base class providing
the interface and generic functions, and the window-system-specific
part providing the actual implementation.

@@

AGL and glX use an event handler singleton, whereas WGL uses one event
handler per window. Pipe event handling is only used for glX, where one
\textsf{Display} connection is created and used to subscribe to window
events.

@@

For the WGL and AGL window systems, it is required to manually receive and
dispatch (`pump') events. On WGL, this has to happen on each thread with
windows, whereas on AGL it has to happen only on the main thread. By
default, Equalizer pumps these events automatically for the application.

@@

Events sent using \textsf{Config::sendEvent} are queued in the
application thread. After a frame has been finished,
\textsf{Config::finishFrame} calls \textsf{Config::handleEvents}. The
default implementation of this method provides non-blocking event
processing, that is, it calls \textsf{Config::handleEvent} for each
queued event. By overriding \textsf{handleEvents}, event-driven
execution can be implemented.

@@

The main idea for direct send is to parallelize the costly recomposition
for database (sort-last) decomposition. With each additional source
channel, the amount of pixel data to be composited grows linearly. When
using the simple approach of compositing all frames on the destination
channel, this channel quickly becomes the bottleneck in the
system. Direct send distributes this workload evenly across all source
channels, and thereby keeps the compositing work per channel constant.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\small\label{fDirectSend}Parallel Direct Send Compositing}}
\end{wrapfigure}
The following operations have to be possible in order to perform this
algorithm:
\begin{itemize}
\item Selection of color and/or depth frame buffer attachments
\item Restricting the read back area to a part of the rendered area
\item Positioning the pixel data correctly on the receiving channels
\end{itemize}

>>>>

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.1.4 for Equalizer 0.4+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.0.1   & Nov 06, 2007 & Fixed Typo\\
    1.1     & Nov 12, 2007 & Adapted distributed objects\\
    1.1.1   & Nov 16, 2007 & Minor improvements based on lulu.com print\\
    1.1.2   & Nov 19, 2007 & Minor improvements based on feedback\\
    1.1.3   & Nov 23, 2007 & Add OpenGL extension handling\\
    1.1.4   & Nov 26, 2007 & Changes caused by removal of glX event thread\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

In general, events are received and processed by the pipe thread a
window belongs to. An exception to this rule is AGL, where all events
are dispatched from the main node thread. WGL and GLX receive and
process the events from the pipe threads that created the
windows. Whenever the term \textbf{event thread} is used, it refers to
the thread receiving the event, i.e., the pipe thread for WGL and GLX,
and the main thread for AGL.

@@

An event handler consists of two parts: the generic base class providing
the interface and generic functions, and the window-system-specific
part providing the actual implementation.

@@

AGL uses an event handler singleton, GLX uses one event handler per pipe
and WGL uses one event handler per window.

@@

In order to dispatch the events, Equalizer 'pumps' the native events. On
WGL and GLX, this has happens on each thread with windows, whereas on
AGL it has to happen only on the main thread. By default, Equalizer
pumps these events automatically for the application in-between
executing task methods.

@@

Events send using \textsf{Config::sendEvent} are queued in the
application thread. After a frame has been finished,
\textsf{Config::finishFrame} calls \textsf{Config::handleEvents}. The
default implementation of this method provides non-blocking event
processing, that is, it calls \textsf{Config::handleEvent} for each
queued event. By overriding \textsf{handleEvents}, event-driven
execution can be implemented.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\small\label{fDirectSend}Parallel Direct Send Compositing}}
\end{wrapfigure}
The main idea behind direct send is to parallelize the costly
recomposition for database (sort-last) decomposition. With each
additional source channel, the amount of pixel data to be composited
grows linearly. When using the simple approach of compositing all frames
on the destination channel, this channel quickly becomes the bottleneck
in the system. Direct send distributes this workload evenly across all
source channels, and thereby keeps the compositing work per channel
constant.

@@

The following operations have to be possible in order to perform this
algorithm:
\begin{itemize}
\item Selection of color and/or depth frame buffer attachments
\item Restricting the read back area to a part of the rendered area
\item Positioning the pixel data correctly on the receiving channels
\end{itemize}
